---
title: "Tidyverse"
output: html_document
---
```{r setup, include=FALSE, cache=FALSE}
library(knitr)
options(digits = 3, scipen = 999)
opts_chunk$set(fig.align='center',
               echo = TRUE, message = FALSE,
               warning = FALSE,
               eval = TRUE, error = TRUE)
```

## Setup

Install and load tidyverse packages

```{r}
# install.packages("tidyverse") # uncomment if you haven't done this yet
library(tidyverse)
```

Set the working directory to `labs` by clicking on the Files pane to the right, navigating to the `labs` folder, and then clicking "More..." "Set As Working Directory".


# What is the "tidyverse"?

> - "an opinionated collection of R packages designed for data science. All packages share an underlying philosophy and common APIs."

> - formerly referred to as the "hadleyverse" for Hadley Wickham

> - packages are strongly associated with RStudio, but not exclusively 

# Core tidyverse packages

- ggplot2
- dplyr 
- tidyr
- readr
- purrr
- tibble
- stringr
- forcats

# 10 differences between the tidyverse and base R


```{r, message = FALSE, echo = FALSE, fig.height = 4}
library(tidyverse)
tidyswiss <- swiss %>% rownames_to_column("Province") %>%
  gather(key, value, -Province)
ggplot(tidyswiss, aes(x = key, y = value)) + geom_boxplot() +
  coord_flip()
```


```{r, fig.height = 5, echo = FALSE}
oldpar <- par(mar = c(4, 6, 2, 2))
boxplot(swiss,
        horizontal = TRUE,
        cex.axis = .8,
        las = 1)
par(oldpar)
```


# 1. Base R

Base: 
```{r, eval = FALSE}
barplot(1:5, horiz = TRUE)
boxplot(1:5, horizontal = TRUE)
```

# 1. Tidyverse is ...  more consistent 

Tidyverse:
```{r, eval = FALSE}
ggplot(...) + geom_bar() + coord_flip()
ggplot(...) + geom_boxplot() + coord_flip()
```


# 2. Base R

Base:
```{r}
df <- data.frame(x = 1:4, y = 1:2)
df
```

# 2. Tidyverse  ...  fails faster

Tidyverse:
```{r}
library(tibble)
df <- tibble(x = 1:4, y = 1:2)
```


# 3. Base R

Base:
```{r}
df <- read.csv("../data/animals.csv")
df
str(df)
```

# 3. Tidyverse ... avoids factors

Tidyverse:
```{r}
tib <- read_csv("../data/animals.csv")
str(tib)
```



# 4. Base R

Base:


```{r}
x <- cbind(df[,1], df[,2]) 
class(x)
```

```{r}
x
```

# 4. Tidyverse ... is more predictable

Tidyverse:


```{r}
x <- bind_cols(tib[,1], tib[,2]) 
class(x)
```

```{r}
x
```


# 4. Base R

Base:

```{r}
class(df)
class(df[,1])
```

# 4. Tidyverse is ... more predictable

Tidyverse:

```{r subsettingtv}
class(tib)
class(tib[,1])
```


# 5. Base R 

Base:
```{r}
# p. 115, Modern Applied Statistics with S-Plus (1999)
x <- rt(250, 9)
qqnorm(x); qqline(x)
```

Source: Venables and Ripley, *Modern Applied Statistics with S-Plus* (1999), p. 115.

(btw 1999 is "new"...)

# 5. Tidyverse is ...  still evolving

Tidyverse:
```{r, warning=TRUE, message = TRUE}
df <- iris %>% dplyr::add_rownames()
```



# 6. Base R

Base:

```{r}
df <- data.frame(x = rnorm(100))
hist(df$x)
```


# 7. Base R

Base:

```{r}
x <- rnorm(100)
hist(x)
```

# 7. Tidyverse ... avoids vectors

Tidyverse:

```{r}
library(ggplot2)
ggplot(x, aes(x)) + geom_histogram()
```

```{r}
ggplot(data.frame(x), aes(x)) + geom_histogram()
```



# 8. Tidyverse is ... more talkative

Tidyverse:
```{r, message = TRUE}
library(ggplot2)
ggplot(df, aes(x)) + geom_histogram()
```

https://twitter.com/hadleywickham/status/952259891342794752



# 8. Tidyverse is .. more coordinated across tasks

Source: RStudio, https://github.com/rstudio/meetup_roadshow/blob/master/2017%20Meetup%20Roadshow.pptx


# 9. Tidyverse is ... easier for beginners

Source: "How `dplyr` replaced my most common R idioms"

http://www.onthelambda.com/2014/02/10/how-dplyr-replaced-my-most-common-r-idioms/

(highly recommended!)

# 10. Tidyverse ... is more collaborative

https://twitter.com/jtrnyc/status/954148122724392960

Source:  https://www.stickermule.com/user/1070448958/stickers
