---
title: "Subsetting data"
author: 
output: html_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE,
                      message = FALSE)
```

## Setup

Install and load tidyverse packages

```{r}
# install.packages("tidyverse") # uncomment if you haven't done this yet
library(tidyverse)
```

Set the working directory to `labs` by clicking on the Files pane to the right, navigating to the `labs` folder, and then clicking "More..." "Set As Working Directory".

Read the data:

```{r}
df <- read_csv("../data/countries2012.csv")
```

## Subsetting rows

a.k.a. extracting cases

### `filter()` (by logical test)

```{r}
df %>% 
  filter(GDP > 100000)
```


```{r}
df %>% 
  filter(COUNTRY == "Canada")
```

```{r}
df %>% 
  filter(CHMORT > TFR)
```

Use functions from the `stringr` package (included with `library(tidyverse)` to work with character data).

```{r}
df %>% 
  filter(str_starts(COUNTRY, "H"))
```

```{r}
df %>% 
  filter(str_length(COUNTRY) == 20)
```

```{r}
df %>%
  filter(str_detect(COUNTRY, " and "))
```


The following four examples produce equivalent results:

```{r}
df %>% 
  filter(GDP >= 40000 & GDP <= 41000)
```


```{r}
df %>% 
  filter(GDP >= 40000, GDP <= 41000)
```


```{r}
df %>% 
  filter(between(GDP, 40000, 41000))
```

```{r}
df %>% 
  filter(GDP >= 40000) %>% 
  filter(GDP <= 41000)
```

### `slice()` (by row number)

```{r}
df %>% 
  slice(1:4)
```

### `sample_n()` (take a random sample of rows)

Run this multiple times:

```{r}
df %>% 
  sample_n(5)
```

## Subsetting columns

a.k.a. extracting variables

Setup: let's work with a smaller version of the data

```{r}
set.seed(12)
random5 <- df %>% 
  sample_n(5)
```

### `select` columns

```{r}
random5 %>% 
  select(COUNTRY)
```

```{r}
random5 %>% 
  select(COUNTRY, GDP)
```

```{r}
random5 %>% 
  select(-GDP)
```

Select consecutive columns:

```{r}
random5 %>%
  select(COUNTRY:GDP)
```


```{r}
random5 %>% 
  select(starts_with("C"))
```

(Note that fucntions such as `starts_with()` only work for selecting columns, not filtering rows.)

Select by column number:

```{r}
random5 %>% 
  select(1:3)
```

... or just use base R:

```{r}
random5[,1:3]
```

```{r}
random5 %>% 
  select(1, 2, 4)
```

or base R:
```{r}
random5 %>% 
  select(c(1, 2, 4))
```

Specify the columns you don't want:

```{r}
random5 %>% 
  select(-2, -4)
```

## Reordering rows

### `arrange()` rows

a.k.a. arranging cases

Note: row order doesn't matter except for creating tables / view data.

```{r}
random5 %>%
  arrange(COUNTRY)
```

```{r}
random5 %>% 
  arrange(desc(TFR))
```

```{r}
random5 %>% 
  select(COUNTRY, GDP) %>% 
  arrange(desc(GDP))
```

## Reordering columns

### `select()` to reorder columns


```{r}
random5 %>% 
  select(COUNTRY, CHMORT, GDP)
```

