---
title: "Transforming data"
author: 
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE,
                      message = FALSE)
```

Read the data:

```{r}
library(tidyverse)
df <- read_csv("../data/countries2012.csv")
set.seed(12)
random5 <- df %>% 
  sample_n(5)
```


## Tranforming column data

### `rename()` columns



```{r}
random5 %>% 
  rename(child_mort = CHMORT)
```

or rename within a call to `select()`:

```{r}
random5 %>% 
  select(country = COUNTRY,
         child_mort = CHMORT,
         fertility = TFR)
```

(Note: in the above example, we did three things: select, rename, and reorder columns.)

### `mutate()` to alter data within a column

In these examples, the same transformation is applied to all rows:

```{r}
random5 %>% 
  mutate(TFR = round(TFR, 1))
```

```{r}
random5 %>% 
  mutate(CONTINENT = str_sub(CONTINENT, 1, 3))
```

```{r}
random5 %>% 
  mutate(CONTINENT = factor(CONTINENT))
```

### `mutate()` to create a new column

Yes... the same function. If the column name doesn't exist, a new one will be created.

```{r}
random5 %>% 
  mutate(TFRROUND = round(TFR))
```

### `mutate()` to convert continuous to categorical data

`if_else()` format: test, if yes, if no

```{r}
df2 <- df %>% 
  mutate(gdp_group = if_else(GDP > mean(GDP), "above average GDP", "below average GDP"))

df2 %>% 
  ggplot(aes(TFR, LIFEEXP)) + 
  geom_point() + 
  facet_wrap(~gdp_group)
```

```{r}
df2 %>% 
  ggplot(aes(TFR, LIFEEXP, color = gdp_group)) + 
  geom_point() 
```

Digression: What is the outlier?

```{r}
df %>% 
  filter(GDP > mean(GDP), TFR > 4)
```

```{r}
df %>% mutate(tfr_group = 
  case_when(
    TFR < 2 ~ "low",
    TFR < 4 ~ "average",
    TFR < 10 ~ "high"
  )
)
```

Following the help instructions to start with the most specific:

```{r}
df %>% mutate(tfr_group = 
  case_when(
    TFR < 2 ~ "low",
    TFR < 4 ~ "average",
    TRUE ~ "high"            # anything else
  )
)
```

Clearer, safer code, imho:

```{r}
df %>% mutate(tfr_group = 
  case_when(
    TFR > 0 & TFR < 2 ~ "low",
    TFR >= 2 & TFR < 4 ~ "average",
    TFR >= 4 ~ "high" 
  )
)
```

Or create a function:

```{r}
tfr_level <- function(x) {
    case_when(
    x > 0 & x < 2 ~ "low",
    x >= 2 & x < 4 ~ "average",
    x >= 4 ~ "high" 
  )
}

df %>% 
  mutate(tfr_group = tfr_level(TFR)) %>% 
  mutate(tfr_group = fct_relevel(tfr_group, "low")) %>% 
  ggplot(aes(CHMORT, LIFEEXP, color = tfr_group)) +
  geom_point()
```


