---
title: "Summarizing data"
author: 
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE,
                      message = FALSE)
```

## Functions

`group_by()`
`summarize()`
`pull()`


## Whole columns

(not using `dplyr`)

Read the data (source: `GDAdata` package):

```{r}
library(tidyverse)
ski <- read.csv("../data/SpeedSki.csv")
head(ski)
summary(ski)
```


From `skimr` package:

```{r}
#install.packages("skimr")
library(skimr)
skim(ski)
```


## `group_by()` + `summarize()` to summarize by factor level

### One group

```{r}
ski %>% 
  group_by(Nation) %>% 
  summarize(MeanSpeed = mean(Speed)) %>% 
  arrange(desc(MeanSpeed))
```

### Multiple groups

```{r}
ski %>% 
  group_by(Nation, Event) %>% 
  summarize(MeanSpeed = mean(Speed)) %>% 
  arrange(desc(MeanSpeed))
```

### Subset by results of `group_by() / summarize()`

```{r}
# find nations with the highest numbers of skiers
top5nations <- ski %>% 
  group_by(Nation) %>%   # group by Nation
  summarize(count = n()) %>% # count the skiers
  arrange(desc(count)) %>%  # reorder the rows
  slice(1:5) %>%  # take the top five
  pull(Nation) # convert to a vector
top5nations
```

```{r}
top5ski <- ski %>% 
  filter(Nation %in% top5nations)
```

### Multiple groups, multiple summary statistics

```{r}
top5ski %>% 
  group_by(Nation, Event, Sex) %>% 
  summarize(MeanSpeed = mean(Speed), Count = n())
```

### Add missing combinations

```{r}
top5ski %>% 
  group_by(Nation, Event, Sex) %>% 
  summarize(MeanSpeed = mean(Speed), Count = n()) %>% 
  ungroup() %>% 
  complete(Nation, Event)

```

### Treating integers as factors

```{r}
top5ski <- top5ski %>% 
  rename(Runs = no.of.runs)
  
top5ski %>%   
  group_by(Event, Runs) %>% 
  summarize(Count = n())
```


```{r}
top5ski %>%   
  group_by(Event, Runs) %>% 
  summarize(Count = n()) %>% 
  ggplot(aes(Runs, Count)) +
  geom_col() +
  facet_wrap(~Event)
```

## Exercise

Read in `NYCdogs.csv`

```{r}
dogs <- read_csv("../data/NYCdogs.csv", na = "n/a")
```

1. Write the code to produce this output:

```
# A tibble: 10 x 3
# Groups:   borough [5]
   borough       spayed_or_neutered count
   <chr>         <chr>              <int>
 1 Bronx         No                  3536
 2 Bronx         Yes                 5757
 3 Brooklyn      No                  5085
 4 Brooklyn      Yes                14248
 5 Manhattan     No                  3866
 6 Manhattan     Yes                22163
 7 Queens        No                  4098
 8 Queens        Yes                13408
 9 Staten Island No                  1455
10 Staten Island Yes                 7926
```

2. Write the code to produce this output:

```
# A tibble: 10 x 4
# Groups:   borough [5]
   borough       spayed_or_neutered count percent
   <chr>         <chr>              <int>   <dbl>
 1 Bronx         No                  3536   0.381
 2 Bronx         Yes                 5757   0.619
 3 Brooklyn      No                  5085   0.263
 4 Brooklyn      Yes                14248   0.737
 5 Manhattan     No                  3866   0.149
 6 Manhattan     Yes                22163   0.851
 7 Queens        No                  4098   0.234
 8 Queens        Yes                13408   0.766
 9 Staten Island No                  1455   0.155
10 Staten Island Yes                 7926   0.845
```

3. Write the code to produce this output:

```
# A tibble: 5 x 2  
# Groups:   borough [5]  
  borough       percent_spayed  
  <chr>                  <dbl>  
1 Bronx                  0.619  
2 Brooklyn               0.737  
3 Manhattan              0.851  
4 Queens                 0.766  
5 Staten Island          0.845  
```

