---
title: "Tidying data"
author: Joyce Robbins
output: 
  md_document:
    toc: true
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE,
                      message = FALSE)
```

## Functions

`gather()`
`spread()`

Online tutorial: https://github.com/jtr13/codehelp/blob/master/R/gather.md

## Messy vs. tidy

```{r}
library(tidyverse)
messydata <- tibble(id = c("car1", "car2", "car3"),
                    city = c(19, 20, 29),
                    hwy = c(24, 30, 35))
messydata
```


```{r}
tidydata <- messydata %>% 
  gather(key = "roadtype",
         value = "mpg", 
         -id)
tidydata
```


## Heads up

There are a few problems I've seen with the way in which people use `gather()`:

### Gathering the `id` column

```{r}
messydata %>% 
  gather(key = "roadtype",
         value = "mpg")
```

### Missing *id* column


```{r}
messy2 <- messydata %>% select("city", "hwy")
messy2

messy2 %>% 
  gather(key = "roadtype", 
         value = "mpg")
```

How to fix: add an `id` column:


```{r}
messy2 %>% rownames_to_column("id") %>% 
  gather(key, value, -id)
```

### Leaving out `key` and `value`


```{r, echo = TRUE, error = TRUE}
messydata %>% 
  gather(-id)
```


```{r, echo = TRUE}
messydata %>% 
  gather(key, value, -id)
```

## Exercise

Tidy the `painters` dataset. It is available in the **MASS** package as well as in the `data` folder.

```{r}
library(MASS)
head(painters)
```

## `gather() + spread()`

Likert data: 25 people and their responses to Q1-Q10

```{r}
df <- read_csv("../data/likert.csv")
df
```

# Number of 1, 2, 3, 4, 5s per person

```{r}
df %>%
  gather(key = "Question", value = "Response", -Name) %>% 
  group_by(Name, Response) %>% 
  summarize(Count = n()) %>% 
  mutate(Response = paste0("Count", Response)) %>% # so column names aren't numbers
  spread(key = "Response", value = "Count")
```


